---
title: "CEMA 0928: Statistics in the Real World"
subtitle: "Constructing Confidence Intervals"
author: "Anthony Scotina"
date: 
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: my-theme.css
    nature:
      highlightLines: true
---

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_accent(base_color = "#43418A")
```

```{r, include = FALSE}
library(tidyverse)
library(moderndive)
library(infer)
```

# Needed Packages

```{r}
library(tidyverse)
library(moderndive)
library(infer)
```

---

# The Percentile Method

The only thing you need to do here is to use the `get_ci()` function from the `infer` package:

```{r}
# From last time...
bootstrap_distribution = pennies_sample %>% 
  specify(response = year) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "mean")
```

```{r, comment = ""}
percentile_ci = bootstrap_distribution %>% 
  get_ci(level = 0.95, type = "percentile") 
# 0.95 and "percentile" are default values
percentile_ci
```

---

# The Percentile Method

The only thing you need to do here is to use the `get_ci()` function from the `infer` package:
```{r, comment = ""}
percentile_ci = bootstrap_distribution %>% 
  get_ci(level = 0.95, type = "percentile") 
# 0.95 and "percentile" are default values
percentile_ci
```

The **percentile method** gives us the 2.5th and the 97.5th *percentiles* of the bootstrap distribution. 

- Our range of plausible values for the mean year on US pennies in 2019 is between 1991 years and 2000 years, with **95% confidence**. 

---

# Visualizing the CI

A cool thing you can do in R is to use the `visualize()` function to plot the *confidence interval* on top of the bootstrap distribution histogram. 
- Run the following:

```{r, eval = FALSE, warning = FALSE}
bootstrap_distribution %>%
  visualize(endpoints = percentile_ci, direction = "between")
```

---

# The Standard Error Method

Our **bootstrap distribution** is close to symmetric and bell-shaped, or **normally distributed**. 

If this is the case, we can use a shortcut formula to calculate the lower and upper endpoints of the confidence interval: 
$$\bar{x}\pm (multiplier)\cdot SE$$
where $\bar{x}$ is the **sample statistic** (`x_bar`, obtained earlier), and $SE$ stands for **standard error**, or the *standard deviation of the bootstrap distribution*.
- For a 95% confidence interval, the value of *multiplier* is **1.96**. 

---

# The Standard Error Method

As with the **percentile method**, we use the `get_ci()` function here. But there are a few slight differences:
```{r, comment = ""}
x_bar = pennies_sample %>%
  summarize(stat = mean(year))
standard_error_ci = bootstrap_distribution %>% 
  get_ci(type = "se", point_estimate = x_bar)
standard_error_ci
```

- We have to specify that we want a CI using the **standard error method** (using `type = "se"`). 
- We also have to provide the **point estimate**, which you should calculate beforehand using `summarize()`. 

---

# The Standard Error Method

As with the **percentile method**, we use the `get_ci()` function here. But there are a few slight differences:
```{r, comment = ""}
standard_error_ci = bootstrap_distribution %>% 
  get_ci(type = "se", point_estimate = x_bar)
standard_error_ci
```

- Our range of plausible values for the mean year on US pennies in 2019 is between 1991 years and 2000 years, with **95% confidence**. 
    - This is identical to the CI obtained using the **percentile method**!
    
- **Note**: The CIs *won't always* be identical! To use the Standard Error Method, you **must** check to see that the bootstrap distribution is *bell-shaped*. 