---
title: "CEMA 0907: Statistics in the Real World"
subtitle: "Interpreting Confidence Intervals"
author: "Anthony Scotina"
date: 
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: my-theme.css
    nature:
      highlightLines: true
      countIncrementalSlides: false
---

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_accent(base_color = "#43418A")
```

```{r, include = FALSE}
library(tidyverse)
library(moderndive)
library(infer)
```

# Needed Packages

```{r}
library(tidyverse)
library(moderndive)
library(infer)
```

---

class: center, middle

# Interpreting the Confidence Interval

---

# Interpreting the Confidence Interval

We have done a lot of calculations in this module, but *what does it all mean*?

The purpose of confidence intervals is to provide a range of plausible values for an *unknown* parameter (in this case, the population mean $\mu$). 
- Using the **percentile method**, one such interval is $[1991, 2000]$. 

--

Let's modify `pennies_sample` to **simulate** a population of US pennies.
```{r, comment = ""}
pennies_pop = pennies_sample %>%
  sample_n(10000, replace = TRUE)

mean(pennies_pop$year)
```

--

Thus, $\mu=1995$ (approx.) which *falls in the confidence interval*. 

---

# Using a new `pennies_sample`

We can use the `infer` pipeline to re-do the entire set of exercises today, but using a *new* sample of pennies from the `pennies` data frame:
```{r, echo = FALSE}
set.seed(45)
```

```{r, message = FALSE}
percentile_ci2 = pennies_pop %>%
  rep_sample_n(size = 40, reps = 1) %>%
  specify(response = year) %>% 
  generate(reps = 1000) %>% 
  calculate(stat = "mean") %>% 
  get_ci()
percentile_ci2
```

This new confidence interval *also* does not contain $\mu$ (1990). 

---

# Calculating Many CIs

We can repeat this process 100 times (don't worry about the code), and generate 100 95% confidence intervals:

.center[
```{r, echo = FALSE, dpi = 300, warning = FALSE, out.width = "60%"}
set.seed(12)
  
# Function to run infer pipeline
bootstrap_pipeline <- function(sample_data){
  sample_data %>% 
    specify(formula = year ~ NULL) %>% 
    generate(reps = 100, type = "bootstrap") %>% 
    calculate(stat = "mean")
}

set.seed(12)
  
# Compute nested data frame with sampled data, sample proportions, all 
# bootstrap replicates, and se_ci
pennies_se_cis <- pennies_pop %>% 
  rep_sample_n(size = 50, reps = 100, replace = FALSE) %>% 
  group_by(replicate) %>% 
  nest() %>% 
  mutate(sample_mean = map_dbl(data, ~mean(.x$year))) %>%
  # run infer pipeline on each nested tibble to generated bootstrap replicates
  mutate(bootstraps = map(data, bootstrap_pipeline)) %>% 
  group_by(replicate) %>% 
  # Compute 95% se CI's for each nested element
  mutate(se_ci = map(bootstraps, get_ci, type = "se", level = 0.95,
                     point_estimate = sample_mean))
  
# Identify if confidence interval captured true p
se_cis <- pennies_se_cis %>% 
  unnest(se_ci) %>% 
  mutate(captured = lower <= 1995.48 & 1995.48 <= upper)
# Plot them!
ggplot(se_cis) +
  geom_segment(aes(
    y = replicate, yend = replicate, x = lower, xend = upper, 
    alpha = factor(captured, levels = c("TRUE", "FALSE"))
  )) +
  geom_point(
    aes(
      x = sample_mean, y = replicate,
      alpha = factor(captured, levels = c("TRUE", "FALSE"))
    ), 
    show.legend = FALSE, size = 1) +
  labs(x = expression("Average Year"), y = "Confidence Interval Number", 
       alpha = "Captured") +
  geom_vline(xintercept = 1995.48, color = "red") + 
  #coord_cartesian(xlim = c(0.1, 0.7)) + 
  theme_light() + 
  theme(panel.grid.major.y = element_blank(), panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())
```
]

---

# Calculating Many CIs

Using 95% as our **confidence level**, *approximately* 95% of the CIs contained the population mean $\mu=1995$, while 8 did not. 

**What does this mean?**
- The procedure we have used to generate confidence intervals is "*95% reliable*" in that we can expect it to include the true population parameter *approximately* 95% of the time *if the process is repeated*.

---

# Back to our example...

**What is a precise interpretation of a confidence interval?**

Recall our *original 95% confidence interval* using the **percentile method**: $[1991, 2000]$. 

**Interpretation**: We are **95% confident** that the average year on US pennies in 2019 is between 1991 and 2000, using the **percentile method**. 

---

# Practice

Using the `diamonds` dataset, calculate a 99% confidence interval for the average price of diamonds. 

--

*Solution*

```{r, comment = "", message = FALSE}
diamonds_ci = diamonds %>%
  specify(response = price) %>%
  generate(reps = 1000) %>%
  calculate(stat = "mean") %>%
  get_ci(level = 0.99)
diamonds_ci
```


---

# Practice

Now, take a random sample of 30 diamonds from the larger sample, and calculate a 99% confidence interval for the average price of diamonds. Use either the percentile or the standard error method. How does this compare to the previous confidence interval?

--

*Solution*

```{r, comment = "", message = FALSE}
diamonds_ci_30 = diamonds %>%
  rep_sample_n(size = 30, reps = 1) %>%
  specify(response = price) %>%
  generate(reps = 1000) %>%
  calculate(stat = "mean") %>%
  get_ci(level = 0.99)
diamonds_ci_30
```

---

# Practice

Because `price` is heavily right-skewed (construct a histogram to see!), calculate a 95% confidence interval for the *median* diamond price. 

--

*Solution*

```{r, comment = "", message = FALSE}
diamonds_ci_median = diamonds %>%
  specify(response = price) %>%
  generate(reps = 1000) %>%
  calculate(stat = "median") %>%
  get_ci()
diamonds_ci_median
```