---
title: "CEMA 0928: Statistics in the Real World"
subtitle: "Scatterplots/Linegraphs"
author: "Anthony Scotina"
date: 
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: my-theme.css
    nature:
      countIncrementalSlides: false
      highlightLines: true
---

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_accent(base_color = "#43418A")
```

```{r, include = FALSE}
library(tidyverse)
library(mdsr)
library(babynames)
library(ggsci)
library(nycflights13)
library(ggthemes)
```

# Needed Packages

```{r, eval = FALSE, message = FALSE, warning = FALSE}
library(nycflights13)
library(tidyverse) # Includes ggplot2
library(mdsr)
```

**Install any packages that you haven't installed yet!** 

---

class: middle, center

# Five Named Graphs: The 5NG

---

# 5NG#1: Scatterplots

**Scatterplots** are one of the most *effective* forms of data visualization. 

- In their simplest forms, they allow one to display the relationship between *two numerical variables*. 

--

Better yet, we've already seen how to make scatterplots!

```{r, eval = FALSE}
ggplot(data = CIACountries, mapping = aes(x = educ, y = gdp)) + 
  geom_point()
```

---

# Scatterplots

In general, we use `geom_point()` to construct a scatterplot, where we can specify *additional aesthetics* (e.g., `color`, `size`) either within `geom_point()`, or *globally* within `ggplot()`. 

We saw this earlier:
```{r, eval = FALSE}
ggplot(data = CIACountries, aes(x = educ, y = gdp, color = net_users)) + 
  geom_point()
```

- Additionally, add `geom_smooth()` to add a smooth trend line to the plot:

```{r, eval = FALSE}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) + 
  geom_point() + 
  geom_smooth()
```

---

# Scatterplots

Add `method = "lm"` and `se = FALSE` to `geom_smooth()` to (i) change the line to a *linear regression line* and (ii) remove the *confidence bands*. 
```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "32%"}
ggplot(data = CIACountries, aes(x = educ, y = gdp)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Education spending (% of GDP)", y = "GDP per capita ($/person)")
```

---

# Scatterplots by group

By adding the `color = net_users` *aesthetic* to `ggplot()`, different colors will be used for different levels of `net_users`, *and* we will also obtain different trend lines. 

```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "35%"}
ggplot(data = CIACountries, aes(x = educ, y = gdp, color = net_users)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Education spending (% of GDP)", y = "GDP per capita ($/person)", 
       color = "Fraction of Internet users")
```

---

# 5NG#2: Line Graphs (Time Series)

**Line Graphs** are similar to scatterplots. They show the relationship between two *numerical* variables. 

- However, linegraphs are useful when the `x` variable is *sequential*, or *ordered*. 

- One of the most common ordered numerical variables is time. 

For example, let's take a look at the `weather` dataset in the `nycflights13` package. Specifically, let's look at **November** *hourly temperature* at the **JFK airport** in New York:

```{r}
jfk_nov = weather %>%
  filter(origin == "JFK" & month == 11)
```

(Don't worry about the syntax yet. Just copy and run!)

---

# Line Graphs

```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "50%"}
ggplot(jfk_nov, aes(x = time_hour, y = temp)) + 
  geom_line()
```

---

# Practice

Using the `diamonds` dataset in the `ggplot2` package (and from a previous example), construct a scatterplot that displays the relationship between `carat` and `price`. 

- Add the `color` aesthetic for the `cut` variable (five categories). 

**Note**: This is a larger dataset than those we've used so far (> 50,000 rows), so run this code to take a sample of 1,000 diamonds if it otherwise takes a while to run. 

```{r, eval = FALSE}
diamonds = diamonds %>%
  sample_n(1000)
```

(Solution on next slide)

---

# Solution

```{r, out.width = "45%", dpi = 300}
diamonds = diamonds %>%
  sample_n(1000)

ggplot(diamonds, aes(x = carat, y = price, color = cut)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)
```



