---
title: "CEMA 0928: Statistics in the Real World"
subtitle: "Histograms/Boxplots"
author: "Anthony Scotina"
date: 
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: my-theme.css
    nature:
      countIncrementalSlides: false
      highlightLines: true
---

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_accent(base_color = "#43418A")
```

```{r, include = FALSE}
library(tidyverse)
library(mdsr)
library(babynames)
library(ggsci)
library(nycflights13)
library(ggthemes)
```

# Needed Packages

```{r, eval = FALSE, message = FALSE, warning = FALSE}
library(nycflights13)
library(tidyverse) # Includes ggplot2
library(mdsr)
library(ggridges)
```

**Install any packages that you haven't installed yet!** 

---

# 5NG#3: Histograms

**Histograms** provide a visualization of the *distribution* of a single *numerical* variable. 

- You need only specific an `x` *aesthetic* in a histogram. 

- By default, the `y` variable is *count*. 

--

Suppose we are interested in the *distribution* of hourly temperature recordings in New York. 

**Histograms** share the following information:

- What is the smallest and largest temperatures, and how often are they observed?
- What is the "center" temperature?
- How are the temperatures spread out?
- What are frequent and infrequent values?
- Is there any skewness?

---

# Histograms

Let's use the `weather` data frame in the `nycflights13` package:

```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "45%"}
ggplot(data = weather, aes(x = temp)) + 
  geom_histogram()
```

---

# Histograms

Before we even discuss the histogram, **always** use `color = "white"` (or some other color) in `geom_histogram()`!!!

```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "40%"}
ggplot(data = weather, aes(x = temp)) + 
  geom_histogram(color = "white")
```

---

# Histograms

.pull-left[
```{r, warning = FALSE, message = FALSE, echo = F, eval = T}
ggplot(data = weather, aes(x = temp)) + 
  geom_histogram(color = "white")
```
]

.pull-right[
What do we notice about the histogram?

- The **middle** temperatures are around 55-60 degrees Fahrenheit. 

- The **range** is from ~10 degrees to ~100 degrees. 

- There are **two prominent peaks** at ~30 degrees and ~70 degrees. 
    - What do you think could account for these two peaks?
    - (Side note: Facets by *month* would be useful here!)
]

---

# Changing the color of histograms

The `color` argument changes the *outline* of each bar in the histogram. 

If you want to change the color of each *bar*, use the `fill` argument. 

- Try running this!
```{r, eval = FALSE, echo = TRUE}
ggplot(data = weather, aes(x = temp)) + 
  geom_histogram(color = "white", fill = "mediumpurple")
```

- R has **many colors**. See [this](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf). 

---

# Changing the bin width of histograms

Note that R usually yells at you when you plot a histogram
```{r, echo = TRUE, fig.show = "hide", warning = FALSE}
ggplot(data = weather, aes(x = temp)) + 
  geom_histogram(color = "white", fill = "mediumpurple")
```

You can use the `binwidth` argument in `geom_histogram()` to specify the **width of the bins** in the histogram. 

- In general, the bin width can change the appearance of a histogram *considerably*, and there usually isn't a single "best" choice. 

--

```{r, eval = FALSE}
ggplot(data = weather, aes(x = temp)) + 
  geom_histogram(color = "white", fill = "mediumpurple", binwidth = 10)
```

Here, each bin contains a ten-point range of temperatures. 

---

# Density Plots

**Density Plots** are an alternate way to visualize the distribution of a single *numerical* variable. 

- Rather than grouping observations into **bins**, a density plot is a single continuous curve over the entire set of observations. 

- The *y*-axis is **density**, and the *area under the curve* totals to 1. 

In `ggplot2`, we can create a density plot using `geom_density()`:

```{r, eval = FALSE}
ggplot(data = weather, aes(x = temp)) + 
  geom_density()
```

---

# Density Plots 

```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "50%"}
ggplot(data = weather, aes(x = temp)) + 
  geom_density()
```

---

# Ridgeline Plots (Density Plots by group)

**Ridgeline plots** consist of multiple, overlapping density plots. These usually resemble a mountain range, hence the name. 

- The `ggridges` package contains an additional **geom**, `geom_density_ridges()`, that allows us to create ridgeline plots. 

--

Using the `weather` data, let's create ridgeline plots that separate temperature by month:

```{r, eval = FALSE}
ggplot(data = weather, aes(x = temp, y = factor(month))) + 
  geom_density_ridges()
```

--

**Note**: `geom_density_ridges()` *requires* the *y* aesthetic to be categorical (i.e., a **factor**). 

- Since the `month` variable is technically *numerical* here, we first convert it to a factor when adding it to the *y* aesthetic. 

---

# Practice

Using the `CIACountries` dataset, create a **ridgeline plot** that separates `gdp` delay by `net_users`. 

```{r, eval = FALSE, echo = FALSE}
ggplot(CIACountries, aes(x = gdp, y = net_users)) + 
  geom_density_ridges()
```

---

# 5NG#4: Boxplots

**Boxplots**, like **histograms** and **density plots**, show the *distribution* of a *numerical* variable. 

- However, boxplots are constructed using information provided by a **five-number summary**. 

**Five-number summary**: Minimum, 1st quartile (25th percentile), Median, 3rd quartile (75th percentile), Maximum

--

In R, a five-number summary of any numerical variable can be found using the `summary()` function:
```{r}
summary(mpg$hwy)
```

(Recall that we can use the `$` operator to view individual variables in a data frame!)

---

# Boxplots

```{r, echo = F, eval = T, out.width = "65%", message = FALSE}
five_number <- tibble(
  hwy = summary(mpg$hwy)[c(1,2,3,5,6)]
)
mpg %>% 
  ggplot(mapping = aes(x = factor(1), y = hwy)) +
  #geom_boxplot() +
  geom_hline(data = five_number, aes(yintercept = hwy), linetype = "dashed") +
  geom_jitter(width = 0.05, height = 0.5, alpha = 0.1) +
  labs(x = "")+ theme(axis.ticks = element_blank(), axis.text.x = element_blank())
```

---

# Boxplots

```{r, echo = F, eval = T, out.width = "65%", message = FALSE}
five_number <- tibble(
  hwy = summary(mpg$hwy)[c(1,2,3,5,6)]
)
mpg %>% 
  ggplot(mapping = aes(x = factor(1), y = hwy)) +
  geom_boxplot() +
  geom_hline(data = five_number, aes(yintercept = hwy), linetype = "dashed") +
  #geom_jitter(width = 0.05, height = 0.5, alpha = 0.1) +
  labs(x = "") + theme(axis.ticks = element_blank(), axis.text.x = element_blank())
```

---

# Boxplots

**How can we make a boxplot?**

Same as before, we just need to change the `geom_` object. 

```{r, echo = TRUE, eval = FALSE}
ggplot(data = mpg, aes(y = hwy)) + 
  geom_boxplot()
```

(Notice we also use the `y` variable here, not the `x`!)

---

# Side-by-side Boxplots

Boxplots are more interesting when you compare several side-by-side. 

- Let's compare *highway miles per gallon* for different *types* of cars. 

--

```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "40%"}
ggplot(data = mpg, aes(x = class, y = hwy)) + 
  geom_boxplot()
```

---

# Violin Plots

**Violin plots** are a sort of *hybrid* of density plots and boxplots:

```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "45%"}
ggplot(data = mpg, aes(x = class, y = hwy)) + 
  geom_violin()
```

---

# Strip Plots

**Strip plots** are a *stripped-down* version of boxplots.
- Use `geom_jitter()` with a **categorical** *x*  and a **numerical** *y* aesthetic.

```{r, message = FALSE, warning = FALSE, dpi = 300, out.width = "35%"}
ggplot(data = mpg, aes(x = class, y = hwy)) + 
  geom_jitter()
```

---

# Note on Categorical x-axis

When using **boxplots**, **violin plots**, or **strip plots**, the *x* and *y* aesthetics are identical. 

- The *x* is **categorical**, and the *y* is **numerical**. 

--

But if you have *too many categories* in your *x* variable, the labels could get jumbled together in the plot. 

- Use `coord_flip()` to fix this!

```{r, eval = FALSE}
ggplot(data = mpg, aes(x = class, y = hwy)) + 
  geom_jitter() + 
  coord_flip()
```

---

# Practice

Using the `diamonds` dataset from the `ggplot2` package, construct an appropriate display for **price** of diamonds. 

- If you think `price` might vary along a *categorical variable* in the dataset (e.g., `cut`, `color`), include it in the plot!
    - [Hint: *violin plots*, *side-by-side boxplots*, *strip plots* could be options if you include a categorical variable.]

```{r, eval = FALSE}
View(diamonds)
?diamonds
```

**Note**: This is a larger dataset than those we've used so far (> 50,000 rows), so run this code to take a sample of 1,000 diamonds if it otherwise takes a while to run. 

```{r, eval = FALSE}
diamonds = diamonds %>%
  sample_n(1000)
```