---
title: "CEMA 0928: Statistics in the Real World"
subtitle: "Multiple Regression (PART 1)"
author: "Anthony Scotina"
date: 
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: my-theme.css
    nature:
      countIncrementalSlides: false
      highlightLines: true
---

<!--
pagedown::chrome_print("~/Dropbox/Teaching/02-Brown Courses/CEMA0928-Statistics in the Real World [ONLINE]/Content/06-Multiple_Regression/06p03-Multiple_Regression.html")
-->

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
mono_accent(base_color = "#4682B4") #3E8A83?
```

```{r, include = FALSE}
library(tidyverse)
library(moderndive)
```

# Needed Packages

```{r}
library(tidyverse)
library(moderndive)
```

---

# Taking care of the outlier...

It is reasonable to suspect that the **outlier** with 33 bedrooms is *not representative* of the population in the same way that the rest of the sample is. 

- Let's remove the outlier to see if `bedrooms` and `price` are more *linearly related*:

```{r}
outlier_index = which(house_prices$bedrooms == 33) #15871
house_prices = 
  house_prices[-outlier_index, ]
```

- This removes the 15,871st row, which contains the outlier, from the data. 

---

# Multiple Regression

**Does bedrooms per house give us the full story?**

```{r, echo = FALSE, out.width = "50%"}
ggplot(data = house_prices, aes(x = bedrooms, y = price)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Bedrooms per home", y = "Price (in $)")
```

---

# Multiple Regression

**Does bedrooms per house give us the full story?**

- *No.*

```{r, echo = FALSE, out.width = "50%"}
ggplot(data = house_prices, aes(x = bedrooms, y = price, color = waterfront)) + 
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Bedrooms per home", y = "Price (in $)")
```

---

# Multiple Regression

**So far**, we have looked at the following *linear regression models* using the `house_prices` data:

- $\widehat{price} = 110316 + 127548*(bedrooms)$
    - For every additional `bedroom`, there is an associated increase of, **on average**, <span>&#36;</span>127,548 on the price of the home. 

--

- $\widehat{price} = 531559 + 1130317*(waterfrontTRUE)$
    - Homes with a *waterfront view* will cost, **on average**, <span>&#36;</span>1,130,317 more than *non-waterfront homes*. 

--

**Why not all at once?**

- $\widehat{price} = b_{0}+b_{1}*(bedrooms)+b_{2}*(waterfrontTRUE)$

---

# Why multiple regression?

It is rare to have a single explanatory (*x*) variable in a study. 

- A *more complex model* can frequently help with **prediction**. 

- The **multiple regression model** is of the form $$\hat{y}=b_{0}+b_{1}x_{1}+b_{2}x_{2}+\cdots+b_{k}x_{k}$$
where there are $k$ explanatory variables. 

---

# One Categorical AND One Explanatory Variable

$$\widehat{price} = b_{0}+b_{1}*(bedrooms)+b_{2}*(waterfrontTRUE)$$

**Question**: Does the *association* between `price` and `bedrooms` change when *simultaneously* considering homes with a *waterfront view*?

--

In this **multiple regression** model, we have:

- A **numerical** response variable (*y*), the price of a home

- *Two* **explanatory variables**:
    - A *numerical explanatory variable*, $x_{1}$, the number of bedrooms per home
    - A *categorical explanatory variable*, $x_{2}$, whether or not the home has a waterfront view
    
---

# Data Visualization

```{r, out.width = "50%"}
ggplot(data = house_prices, aes(x = bedrooms, y = price, color = waterfront)) + 
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Bedrooms per home", y = "Price (in $)")
```

---

# Data Visualization

.pull-left[
```{r, echo = FALSE}
ggplot(data = house_prices, aes(x = bedrooms, y = price, color = waterfront)) + 
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Bedrooms per home", y = "Price (in $)")
```
]

.pull-right[
**Some observations**

- Both slopes are **positive**. 

- But, the slope for *waterfront view homes* is **more positive**. 

In other words, the average house price per bedroom is *higher* for *waterfront view homes*. 
]

---

# Interaction Model

We will quantify the relationship between the response *y* and the *two* explanatory variables using a type of multiple regression model known as an **interaction model**. 

```{r}
lm_price_interaction = lm(price ~ bedrooms + waterfront + 
                            bedrooms*waterfront, 
                          data = house_prices)
get_regression_table(lm_price_interaction)
```

In this **interaction model**, the model formula was not only of the form `y ~ x1 + x2`, but it included an **interaction term**, `x1*x2`. 

---

# Interaction Model

```{r, echo = FALSE}
knitr::kable(get_regression_table(lm_price_interaction)[,-c(4,5)], format = "html")
```

First, recall that *non-waterfront homes* (`waterfront = FALSE`) serves as the **reference** (or *baseline*) group. 
- Therefore, the `intercept` term is the **intercept** *for only the non-waterfront homes*. 

- Similarly, the `bedrooms` term is the **slope** *for only the non-waterfront homes*.

What about the slope and intercept for *waterfront homes*?

---

# Linear Model Equation

```{r, echo = FALSE}
knitr::kable(get_regression_table(lm_price_interaction)[,-c(4,5)], format = "html")
```

We can write the equation of the linear model *with interaction term* as 
\begin{align*}
\widehat{price}&=114143+123862(bedrooms)-236296(waterfrontTRUE)\\
&\ \ \ \ +416652(bedrooms:waterfrontTRUE)
\end{align*}

---

# Interaction Model

\begin{align*}
\widehat{price}&=114143+123862(bedrooms)-236296(waterfrontTRUE)\\
&\ \ \ \ +416652(bedrooms:waterfrontTRUE)
\end{align*}

The value for `waterfrontTRUE` of $-236296$ is *not* the intercept for waterfront homes.
- Rather, it serves as the **offset** in the intercept for waterfront homes *relative to non-waterfront homes*. 

- The *intercept* for *waterfront homes* is `intercept` + `waterfrontTRUE` = $114143-236296$ = $-122153$.

--

**Uhhh wut?**

- I thought waterfront homes were *more expensive*!!!

---

# The Interaction Term

\begin{align*}
\widehat{price}&=114143+123862(bedrooms)-236296(waterfrontTRUE)\\
&\ \ \ \ +416652(bedrooms:waterfrontTRUE)
\end{align*}

The `bedrooms:waterfrontTRUE` is *not* the slope for age for waterfront homes. 
- Rather, it serves as the **offset** in the slope for waterfront homes. 

The *slope* for *waterfront* homes is `bedrooms` + `bedrooms:waterfrontTRUE` = $123862+416652$ = $540514$. 

---

# Putting It All Together

\begin{align*}
\widehat{price}&=114143+123862(bedrooms)-236296(waterfrontTRUE)\\
&\ \ \ \ +416652(bedrooms:waterfrontTRUE)
\end{align*}

Because there are *two levels* of the categorical explanatory variable (`TRUE` and `FALSE`), we can write the regression model as equations for *two separate regression lines*:

- For `waterfront = TRUE`, $$\widehat{price}=-122153+540514(bedrooms)$$
- For `waterfront = FALSE`, $$\widehat{price}=114143+123862(bedrooms)$$

---

# Putting It All Together

.pull-left[
```{r, echo = FALSE}
ggplot(data = house_prices, aes(x = bedrooms, y = price, color = waterfront)) + 
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x = "Bedrooms per home", y = "Price (in $)")
```
]

.pull-right[
- For `waterfront = TRUE`, $$\widehat{price}=-122153+540514(bedrooms)$$
- For `waterfront = FALSE`, $$\widehat{price}=114143+123862(bedrooms)$$
]

---

# Interpreting the Regression Lines

For `waterfront = TRUE`, $$\widehat{price}=-122153+540514(bedrooms)$$

- For **homes with a waterfront view**: For every additional `bedroom`, there is an associated increase of, **on average**, <span>&#36;</span>540,514 on the price of the home. 

--

For `waterfront = FALSE`, $$\widehat{price}=114143+123862(bedrooms)$$

- For **homes without a waterfront view**: For every additional `bedroom`, there is an associated increase of, **on average**, <span>&#36;</span>123,862 on the price of the home. 

--

Thus, our model is suggesting that the number of bedrooms impacts the price of *waterfront homes* **more** than it does for *non-waterfront homes*. 

---

# Summary of Interaction Model

\begin{align*}
\widehat{price}&=114143+123862(bedrooms)-236296(waterfrontTRUE)\\
&\ \ \ \ +416652(bedrooms:waterfrontTRUE)
\end{align*}

- $b_{0}=114143$ is the `intercept` for *non-waterfront homes*.

- $b_{1}=123862$ is the slope for `bedrooms` for *non-waterfront homes*.

- $b_{2}=-236296$ is the **offset** in the intercept for *waterfront homes*. 

- $b_{3}=416652$ is the **offset** in the slope for *waterfront homes*. 

---

# Back to Interaction Effects

Why do we call that additional term the **interaction effect**?

- We say there is an interaction effect if the *associated effect* of one variable *depends* on the value of another variable. In other words, the two variables are “*interacting*” with each other.

--

For a given home in King County, WA, there might be an associated effect of `bedrooms` *by itself*. 

There might also be an associated effect of `waterfront` homes *by itself*. 

- **But** when `bedrooms` and `waterfront` homes are considered *together*, there might be *an additional effect* beyond those present when considering the variables separately. 
